### 2020-07-28 thesis/thesis.tex
[thesis] More on universal algebra

%%%%% WLASNE DODATKOWE PAKIETY
% \usepackage{graphicx,listings,tikz}
\usepackage{syntax}
\usepackage{mathalfa}
\usepackage{textcomp}
\usepackage{stmaryrd}

@@

For the purpose of describing the duality between algebra and coalgebra,
we allow ourselves to wander a bit deeper into category theory.

@@

    \begin{definition}{\mathVar{Signature}}

@@

    \begin{definition}{$\mathVar{\Sigma-terms}$}

@@

    \begin{definition}{$ \mathVar{\Sigma-Equation} $}

@@

        $ X $ $|$ $l = r$

@@

    \begin{definition}{Algebraic Theory}

@@

    \begin{definition}{Interpretation}

@@

    $ {\llbracket op_{i} \rrbracket}_I : P_{i} \times{} {|I|}^{A_{i}} \rightarrow |I| $

    Interpretation may be naturally extended to $\mathVar{\Sigma-terms}$, such
    that a given $\mathVar{\Sigma-term}$ $X$ $|$ $t$ is interpreted by a map

    $ {\llbracket X | t \rrbracket}_I : {|I|}^{k} \rightarrow |I| $

    TODO actually define the map as a projection on variables and
    map composition on compound terms.

    where ${|I|}^{k}$ is an k-ary product in our category, assuming that all of the
    necessary products exist.

@@

    TODO Models, Free models

    \subsection{Free models form monads}

    Maybe not a subsection but just a lemma

    \subsection{Models homomorphisms}

    T-homomorphisms with respect to theory T

    \subsection{Parametrized operations}

>>>>

%%%%% WLASNE DODATKOWE PAKIETY
% \usepackage{graphicx,listings,tikz}
\usepackage{syntax}
\usepackage{mathalfa}
\usepackage{textcomp}
\usepackage{stmaryrd}
\usepackage{tikz-cd}
% Not sure if necessary
\usepackage[utf8]{inputenc}
\usepackage{rotchiffre}

@@

Algebraic effects can be describe via operational means, however,
for the purpose of presenting the duality between algebra and coalgebra,
we allow ourselves to wander a bit deeper into category theory and describe
effects from denotational point of view.

@@

    \begin{definition}

@@

    \begin{definition}

@@

    \begin{definition}

@@

        \begin{center}
        $ X \mid l = r $
        \end{center}

@@

    \begin{definition}

@@

    \begin{definition}

@@

    \begin{center}
        $ {\llbracket op_{i} \rrbracket}_I : P_{i} \times{} {|I|}^{A_{i}} \rightarrow |I| $
    \end{center}
    Interpretation may be naturally extended to $\mathVar{\Sigma-terms}$, such
    that a given $\mathVar{\Sigma-term}$ $X \mid t$ is interpreted by a map
    which sends variables into projections from environment and terms into
    map composition over each subterm.

@@

    \begin{definition}

    A \textit{model M} of an algebraic theory T is an interpretation of the
    signature ${\Sigma_{T}}$ which validates all the equations $\mathcal{E_{T}}$.
    That is, for every equation $X \mid l = r$ the following diagram commutes:

    \begin{center}
    \begin{tikzcd}[row sep=huge, column sep=large]
        {|M|}^k \arrow[rr, '\phi', bend right] \arrow[rr, '\rrbracket{} X|l', bend left] &  & {|M|}
    \end{tikzcd}
    \end{center}
    % TODO: this diagram doesn't work

    \begin{center}
    \begin{tikzcd}
    A \arrow[rd] \arrow[r, '\phi'] & B \\
    & C
    \end{tikzcd}
    \end{center}

    \begin{definition}

    Free F-algebra on an object A (of generators) in $\mathcal{C}$ is meant an algebra

    \begin{center}

    $ \varphi_{A} : F A^{\#} \longrightarrow A $

    \end{center}
    \\
    together with an universal arrow $ \eta_{A} : A \longrightarrow A^{\#} $. Universality means that for every algebra
    $ \beta : F B \longrightarrow B $ and every morphism $ f : A \longrightarrow B $ in $ \mathcal{C}$, there exists a unique homomorphism
    $ \overline{f} : A^{\#} \longrightarrow B $ extending f, i.e.\ a unique morphism of $ \mathcal{C}$ for which the diagram below
    commutes:

    % TODO:
    \begin{center}
    \begin{tikzcd}[row sep=huge, column sep=large]
    % F A^{\#} \arrow[d, "F \overline{f}"] \arrow[r, "\varphi_{A}"] & A^{\#} \arrow[d, "\overline{f}"] & A \arrow[l, "\eta_{A}"'] \arrow[ld, "f"] \\
    % F B \arrow[r, "\beta"']                                       & B                                &
    \end{tikzcd}
    \end{center}

    \begin{definition} \textit{Free model} is just a model which is free algebra.

    \begin{lemma}{Free models form monads}

    \end{lemma}

    \begin{definition}

    Let L, M be models of a theory T. A $ \mathVar{T-homomorphism}$
    $\phi : L \rightarrow M$ is a map such that the following diagram commutes:

    % TODO
